---
title: "highTPM_nonAS_genes"
author: "Anqi Dai"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 3
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = F, warning = F)
```

```{r}
# this script will find the gene name and description for the high TPM & non AS genes
library(kableExtra)
library(tidyverse)
library(biomaRt)
library(knitr)
```

## How many genes

The genes will be the ones that are non AS significant but have high TPM (at least 50).

There are a total of: 

```{r}

# the genes will be the ones that suggest ribosomal biogenesis enrichment when comparing KD to control.
setwd('~/SDE2/analysis/11_DE/scripts/')

genes <- read_table('../../09_mitochondria/output/genes_with_at_least_3_introns_ids_highTPM.txt', col_names = F) %>%
  pull(X1)

length(genes) # there are 787 genes in it

# find the gene name and description for them

mart <- useMart("ensembl", dataset = "hsapiens_gene_ensembl")

# retrieve the feature information of the listed attributes
symbol <- getBM(filters = "ensembl_gene_id",
                attributes = c('ensembl_gene_id',
                               'external_gene_name',
                               'chromosome_name',
                               'start_position',
                               'end_position', 
                               'description'),
                values = genes, 
                mart = mart)


symbol <- symbol %>%
  rename(chr = chromosome_name)

write_csv(symbol, '../output/01_highTPM_non_AS_gene_info.csv')

```

## Result full table 

[TABLE attached: high TPM and non AS genes info](../output/01_highTPM_non_AS_gene_info.csv)

Below is a web representation of the same table. You can directly search on this page.

```{r}
symbol %>%
  kable() %>%
  kable_styling()
```

## A look at ribosome related ones

There are a total of: 

```{r}
symbol %>%
  filter(grepl(pattern = 'ribo', x = description)) %>%
  nrow
```

```{r}
symbol %>%
  filter(grepl(pattern = 'ribo', x = description)) %>%
  kable() %>%
  kable_styling()
```

## The AS genes

Compare to how many genes that are ribosome related but are AS significant.

### How many genes are AS sig pooling all the AS software we used:

```{r}
# load the 8000 genes that are AS significant

# use the detk filter to be the filtered result
detkF <- read_csv('../../08_findingGenes/data/detk_fil_Norm_salmon_cts_replicate1_matrix.csv') %>%
  mutate(Name = str_replace_all(string = Name, pattern = '\\..*$', replacement = '')) 
  
# loading the sig genes
sig_gene <- read_table('../../08_findingGenes/data/union_sig_genes_from_3_analysis.txt', col_names = F)

sig_overlap <- intersect(detkF$Name, sig_gene$X1)


```

```{r}
# retrieve the feature info for those genes

mart <- useMart("ensembl", dataset = "hsapiens_gene_ensembl")
# retrieve the feature information of the listed attributes
symbol <- getBM(filters = "ensembl_gene_id",
                attributes = c('ensembl_gene_id',
                               'external_gene_name',
                               'chromosome_name',
                               'start_position',
                               'end_position', 
                               'description'),
                values = sig_overlap, 
                mart = mart)
symbol <- symbol %>%
  rename(chr = chromosome_name)

symbol %>%
  nrow

write_csv(symbol, '../output/01_all_AS_gene_info.csv')
```

[TABLE attached: all AS sig genes info](../output/01_all_AS_gene_info.csv)

### How many are ribosome related

```{r}
symbol %>%
  filter(grepl(pattern = 'ribosom', x = description)) %>%
  nrow
```

```{r}
symbol %>%
  filter(grepl(pattern = 'ribosom', x = description)) %>%
  kable() %>%
  kable_styling()
```


