---
title: "Encode KD experiments VS SDE2 KD experiment"
author: "Anqi Dai"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 3
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = F, warning = F)
library(tidyverse)
library(kableExtra)
library(ggpubr)
library(RVenn)
```

## Load the data from analyzing the ENCODE data

```{r}
encode_data_path <- '../data/encode_IRfinder_raw_data'
encode_fns <- list.files(encode_data_path, 'txt', full.names = T)
```

## Get the sig events of ENCODE data

```{r}
# the whole sig df
en_sig <- encode_fns %>% 
  map(~ read_tsv(.,comment = '#',col_names = T)  %>%
        mutate(Start = Start + 1) %>%
        mutate(locus = paste(Chr,paste(Start, End, sep = '-'), sep = ':' ) ) %>%
        rename_all(funs(stringr::str_replace_all(., '-', '_') )) %>%
        separate(names(.)[4], into = c('Symbol', 'GeneID', 'Category'), sep = '/') %>%
        mutate(ensembl_gene_id = str_replace_all(string = GeneID, pattern = '\\..*$', replacement = ''),
               padj = p.adjust(p_diff, method = 'BH')) %>%
        filter(! Category == 'known-exon') %>%
        filter(A_IRok != 'MinorIsoform' &  B_IRok != 'MinorIsoform' ) %>%
        filter(padj < 0.05))

# how many sig events are there under FDR < 0.05
encode_sig_df <- en_sig %>% 
  set_names(encode_fns) %>% 
  map_dfc(~ nrow(.)) %>%
  gather(key = 'KD_gene', 
         value = 'num_sig') %>%
  mutate(KD_gene = str_replace(KD_gene, '../data/encode_IRfinder_raw_data/',''),
         KD_gene = str_replace(KD_gene, '_vs_ctrl.txt',''))
```

```{r}
# what are those 8 genes are
library(biomaRt)

mart <- useMart("ensembl", dataset = "hsapiens_gene_ensembl")

# the ensembl_gene_id of the genes that you are interested in 
genes <- encode_sig_df$KD_gene

# retrieve the feature information of the listed attributes
symbol <- getBM(filters = "external_gene_name",
                attributes = c('external_gene_name','description'),
                values = genes, 
                mart = mart)


```


```{r}
# pull the locus column
en_sig_locus <- en_sig %>%
  map(~ pull(.data = ., locus))
```


## Our own sig events of SDE2

```{r}
# our SDE2 data

data_path <- '../../05_IRfinder/data/rawIRfinderResult/'
fns <- list.files(data_path, pattern = 'txt', full.names = T)

SDE2_sig_all <- fns %>%
  map(~ read_tsv(.,comment = '#',col_names = T)  %>%
        mutate(Start = Start + 1) %>%
        mutate(locus = paste(Chr,paste(Start, End, sep = '-'), sep = ':' ) ) %>%
        rename_all(funs(stringr::str_replace_all(., '-', '_') )) %>%
        separate(names(.)[4], into = c('Symbol', 'GeneID', 'Category'), sep = '/') %>%
        mutate(ensembl_gene_id = str_replace_all(string = GeneID, pattern = '\\..*$', replacement = ''),
               padj = p.adjust(p_diff, method = 'BH')) %>%
        filter(! Category == 'known-exon') %>%
        filter(A_IRok != 'MinorIsoform' &  B_IRok != 'MinorIsoform' ) %>%
        filter(padj < 0.05))


# an SDE2_sig_all in df form to have the gene name
# the df that has all the SDE2 sig locus essential info like gene name
SDE2_sig_info <- SDE2_sig_all %>% 
  map_dfr(~ dplyr::select(.data = ., locus, ensembl_gene_id, external_gene_name = Symbol)) %>%
  distinct(locus, .keep_all = T)

SDE2_sig_locus <- SDE2_sig_all %>%
  map(~ pull(.data = ., locus))


# intersect and union of the SDE2 KD
SDE2_intersect <- intersect(SDE2_sig_locus[[1]],
                            SDE2_sig_locus[[2]])


SDE2_union <- union(SDE2_sig_locus[[1]],
                    SDE2_sig_locus[[2]])
```

```{r}
# a reminder of how many events these are
SDE2_num_summary <- list(SDE2_intersect = SDE2_intersect,
     SDE2_union = SDE2_union) %>% 
  imap_dfc(~ length(.)) 
```
 
## Compare the two

Specifically, the sig in SDE2 is the intersection of 14 and 89.

```{r}
# compare each of the encode KD with SDE2 intersect
ret_intersect <- en_sig_locus %>% 
  set_names(encode_sig_df$KD_gene) %>% 
  map(~ intersect(.,SDE2_intersect)) 
```

```{r}
# some set operation to see the relationship among those KD events
intersect_set <- construct(ret_intersect)

```


```{r}
# a df of the specific intersect locus between encode data and our data
# and with the gene info and name
ret_intersect_info_df <- ret_intersect %>% 
  map(~ tibble(locus = .) %>% 
        left_join(SDE2_sig_info, by = 'locus'))  %>% 
  bind_rows(.id = 'column_label')
  

ret_intersect_info_df %>% 
  write_csv('../output/locus_info_for_each_encode_KD_gene_intersect_with_14&89_overlap.csv')
```

**A table will be sent together with this file with name "locus_info_for_each_encode_KD_gene_intersect_with_14&89_overlap.csv", including info summarized in the bottom table of this page and also have gene name.**


```{r}
# these are the events exist in all encode KD experiments and both our 14 and 89
all_existing <- overlap(intersect_set)

SDE2_sig_info %>% 
  filter(locus %in%  all_existing) %>% 
  kable(caption = 'Info for the locus that exist in all encode KD experiments and both our 14 and 89') %>% 
  kable_styling(full_width = F, position = 'left')


ggvenn(intersect_set, slice = c(1, 2, 3 )) +
  labs(title = 'An example Venn diagram showing the relationship\nof sig locus in different ENCODE KD experiments')
# ggvenn(intersect_set, slice = c(4, 5, 6 ))
# ggvenn(intersect_set, slice = c(2, 7, 8 ))
```

```{r}
# the stats summary
intersect_stats <- ret_intersect %>% 
  imap_dfc(~ length(.)) %>% 
  gather(key = 'KD_gene',
         value = 'num') %>% 
  full_join(symbol %>% 
              rename(KD_gene = external_gene_name), by = 'KD_gene') %>% 
  full_join(encode_sig_df, by  = 'KD_gene') %>% 
  mutate(SDE2_intersect_num = length(SDE2_intersect))%>% 
  dplyr::select(KD_gene, 
                num_sig,
                SDE2_intersect_num,
                overlap_num = num,
                description) 

intersect_stats %>% 
  kable(caption = "Summary of the encode gene KD sig IR events' overlap with the SDE2 14 & 89 intersected sig events") %>% 
  kable_styling()

```

#### An explaination of the above table:

+ num_sig: the number of sig events (FDR < 0.05) of each encode KD experiment calculated in the same way that I did my SDE2 calculations
+ SDE2_intersect_num: the number of sig events that exist in both 14 and 89
+ overlap_num: the number of overlapped locus between the previous two columns
+ description: the description of the KD gene

