---
title: "PSI14 and 89"
author: "Anqi Dai"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 3
    code_folding: hide
---

Day 5.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = F, warning = F)
library(tidyverse)
library(ggpubr)
library(kableExtra)  
```

Get the significant results in each tool.

How to get the intersection in each tool:

+ Get the significant results in 14 and 89 respectively
+ Take the intersection 

# Whippet intersection 

```{r}
# data path for whippet
data_path <- '../../03_ASprofile/data/whippetResult'
fns <- list.files(data_path, pattern = 'output.diff', full.names = T)

# read the data and select the columns
whippet_sig <- fns %>%
  map(~ read_tsv(.,col_names = T)  %>%
        mutate(absDeltaPsi = abs(DeltaPsi)) %>%
        filter(absDeltaPsi > 0.1 & Probability > 0.9 & Psi_A > 0.1) %>%
        mutate(tool = 'Whippet') %>%
        dplyr::select(ensembl_gene_id = Gene, 
                      locus = Coord, 
                      type = Type,
                      Psi_A, 
                      Psi_B,
                      tool) %>%
        mutate(ensembl_gene_id = str_replace_all( ensembl_gene_id, '\\..*$', ''))
   )

# find the intersection by inner join
whippet_ol  <- whippet_sig[[1]] %>%
  rename(Psi_14 = Psi_A) %>%
  inner_join(whippet_sig[[2]] %>%
               dplyr::select(Psi_89 = Psi_A, locus), by = 'locus')
        
```

# IRfinder intersection

```{r}
data_path <- '../../05_IRfinder/data/rawIRfinderResult/'
fns <- list.files(data_path, pattern = 'txt', full.names = T)

IRfinder_sig <- fns %>%
  map(~ read_tsv(.,comment = '#',col_names = T)  %>%
        mutate(Start = Start + 1) %>%
        mutate(locus = paste(Chr,paste(Start, End, sep = '-'), sep = ':' ) ) %>%
        rename_all(funs(stringr::str_replace_all(., '-', '_') )) %>%
        separate(names(.)[4], into = c('Symbol', 'GeneID', 'Category'), sep = '/') %>%
        mutate(ensembl_gene_id = str_replace_all(string = GeneID, pattern = '\\..*$', replacement = ''),
               padj = p.adjust(p_diff, method = 'BH')) %>%
        filter(! Category == 'known-exon') %>%
        filter(A_IRok != 'MinorIsoform' &  B_IRok != 'MinorIsoform' ) %>%
        filter(padj < 0.05)  %>%
        mutate(type = "RI",
               tool = 'IRfinder') %>%
        dplyr::select(ensembl_gene_id, 
                      locus,
                      type,
                      Psi_A = A_IRratio, 
                      Psi_B = B_IRratio,
                      tool))

IRfinder_ol <- IRfinder_sig[[1]] %>%
  rename(Psi_14 = Psi_A) %>%
  inner_join(IRfinder_sig[[2]] %>%
               dplyr::select(Psi_89 = Psi_A, locus), by = 'locus')
```

# rMATS intersection

```{r}
# the function of combine all of the types of results in rMATS
# input Name could be: RF14 or RF89.
   
  
read_rMATS_all_types <- function(Name) {

  path <- str_glue('../../06_rMATS/data/rMATS_result/{Name}_control')
  fns <- file.path(path, list.files(path, 'JCEC.txt'))
  
  # 1. A3SS  
  A3SS <- read_tsv(fns[1]) %>% 
    filter(FDR < 0.05) %>%
    rename(start = longExonStart_0base, 
           end = shortES, 
           Psi_A = IncLevel1, 
           Psi_B = IncLevel2) %>%
    mutate(start = start + 1,
           locus = paste(chr, paste(start,end, sep='-'),sep=':'),
           tool = 'rMATS',
           type = 'AA') %>%
    separate(Psi_A, into = c('Psi_A1','Psi_A2','Psi_A3'), sep = ',', convert = T) %>%
    separate(Psi_B, into = c('Psi_B1','Psi_B2','Psi_B3'), sep = ',', convert = T) %>%
    rowwise() %>%
    mutate(Psi_A = mean(c(Psi_A1, Psi_A2, Psi_A3), na.rm = T),
           Psi_B = mean(c(Psi_B1, Psi_B2, Psi_B3), na.rm = T))  %>%
    dplyr::select(ensembl_gene_id = GeneID, locus, type, Psi_A,  Psi_B, tool) %>%
    mutate(ensembl_gene_id = str_replace_all( ensembl_gene_id, '\\..*$', ''))

  
  # 2. A5SS   
  A5SS <- read_tsv(fns[2]) %>%
    filter(FDR < 0.05) %>% 
    rename(start = shortEE, end = longExonEnd, Psi_A = IncLevel1, Psi_B = IncLevel2) %>%
    mutate(locus = paste(chr, paste(start,end, sep='-'),sep=':'),
       tool = 'rMATS',
       type = 'AD') %>%
    separate(Psi_A, into = c('Psi_A1','Psi_A2','Psi_A3'), sep = ',', convert = T) %>%
    separate(Psi_B, into = c('Psi_B1','Psi_B2','Psi_B3'), sep = ',', convert = T) %>%
    rowwise() %>%
    mutate(Psi_A = mean(c(Psi_A1, Psi_A2, Psi_A3), na.rm = T),
           Psi_B = mean(c(Psi_B1, Psi_B2, Psi_B3), na.rm = T))  %>%
    dplyr::select(ensembl_gene_id = GeneID, locus, type, Psi_A,  Psi_B, tool) %>%
    mutate(ensembl_gene_id = str_replace_all( ensembl_gene_id, '\\..*$', ''))
    
  #  RI 
  RI <- read_tsv(fns[4]) %>%
    filter(FDR < 0.05) %>% 
    rename(start = upstreamEE, end = downstreamES, Psi_A = IncLevel1, Psi_B = IncLevel2) %>%
    separate(Psi_A, into = c('Psi_A1','Psi_A2','Psi_A3'), sep = ',', convert = T) %>%
    separate(Psi_B, into = c('Psi_B1','Psi_B2','Psi_B3'), sep = ',', convert = T) %>%
    rowwise() %>%
    mutate(Psi_A = mean(c(Psi_A1, Psi_A2, Psi_A3), na.rm = T),
           Psi_B = mean(c(Psi_B1, Psi_B2, Psi_B3), na.rm = T))  %>%
    mutate(locus = paste(chr, paste(start,end, sep='-'),sep=':'),
           tool = 'rMATS',
           type = 'RI') %>%
    dplyr::select(ensembl_gene_id = GeneID, locus, type, Psi_A,  Psi_B, tool) %>%
    mutate(ensembl_gene_id = str_replace_all( ensembl_gene_id, '\\..*$', ''))

  
  #  SE 
  SE <-  read_tsv(fns[5]) %>%
    filter(FDR < 0.05) %>% 
    rename(start = exonStart_0base, end = exonEnd, Psi_A = IncLevel1, Psi_B = IncLevel2) %>%
    mutate(start = start + 1,
           locus = paste(chr, paste(start,end, sep='-'),sep=':'),
           tool = 'rMATS',
           type = 'CE') %>%
    separate(Psi_A, into = c('Psi_A1','Psi_A2','Psi_A3'), sep = ',', convert = T) %>%
    separate(Psi_B, into = c('Psi_B1','Psi_B2','Psi_B3'), sep = ',', convert = T) %>%
    rowwise() %>%
    mutate(Psi_A = mean(c(Psi_A1, Psi_A2, Psi_A3), na.rm = T),
           Psi_B = mean(c(Psi_B1, Psi_B2, Psi_B3), na.rm = T))  %>%
    dplyr::select(ensembl_gene_id = GeneID, locus, type, Psi_A,  Psi_B, tool) %>%
    mutate(ensembl_gene_id = str_replace_all( ensembl_gene_id, '\\..*$', ''))
  
  #  MXE
  MXE <- read_tsv(fns[3]) %>%
    filter(FDR < 0.05) %>% 
    rename(start = upstreamES, end = downstreamEE, Psi_A = IncLevel1, Psi_B = IncLevel2) %>%
    mutate(locus = paste(chr, paste(start,end, sep='-'),sep=':'),
           tool = 'rMATS',
           type = 'MXE') %>%
    separate(Psi_A, into = c('Psi_A1','Psi_A2','Psi_A3'), sep = ',', convert = T) %>%
    separate(Psi_B, into = c('Psi_B1','Psi_B2','Psi_B3'), sep = ',', convert = T) %>%
    rowwise() %>%
    mutate(Psi_A = mean(c(Psi_A1, Psi_A2, Psi_A3), na.rm = T),
           Psi_B = mean(c(Psi_B1, Psi_B2, Psi_B3), na.rm = T))  %>%
    dplyr::select(ensembl_gene_id = GeneID, locus, type, Psi_A,  Psi_B, tool) %>%
    mutate(ensembl_gene_id = str_replace_all( ensembl_gene_id, '\\..*$', ''))

  
  # combine them together
  sig <- bind_rows(A3SS, A5SS, RI, SE, MXE) %>%
    distinct(locus, .keep_all = TRUE) 
  
  return(sig)
}

```

```{r}
# call the above function
rMATS_sig <- c('RF14','RF89') %>%
  map(~ read_rMATS_all_types(.))
  
# join to find the intersection
rMATS_ol <- rMATS_sig[[1]] %>%
  rename(Psi_14 = Psi_A) %>%
  inner_join(rMATS_sig[[2]] %>% 
               dplyr::select(Psi_89 = Psi_A, locus), by = 'locus')
```

```{r}
# how many of them each
tibble(
  IRfinder_ol = nrow(IRfinder_ol),
  whippet_ol = nrow(whippet_ol),
  rMATS_ol = nrow(rMATS_ol)
) %>%
  kable(caption = 'Number of overlapped sig events between 14 and 89 in each tool') %>%
  kable_styling(full_width = F, position = 'left')
```


# Combine three

+ Combine all the significant results from three tools together, and remove duplicated locus.
+ PSI_KD is the mean of Psi_14 and Psi_89.
+ PSI_KD - PSI_WT = deltaPSI
+ The table is sorted in the descending order of abs(deltaPSI)

```{r}
all_sig <- bind_rows(
  IRfinder_ol,
  whippet_ol,
  rMATS_ol
) %>%
  distinct(locus, .keep_all = TRUE) 
```

```{r}
# calculate the mean of 14 and 89 as the PSI_KD and
# PSI_KD - PSI_B = deltaPSI
ret <- all_sig %>%
  rowwise() %>%
  mutate(PSI_KD = mean(c(Psi_14, Psi_89), na.rm = T)) %>%
  rename(PSI_WT = Psi_B) %>%
  mutate(deltaPSI = PSI_KD - PSI_WT) %>%
  arrange(desc(abs(deltaPSI)))
```


```{r}
# write out the above table . this is the result of sig events not filtered by PSI or deltaPSI
ret %>%
  write_csv('../output/day5_sig_events_all_tools_not_filtered_by_PSI.csv')
```

