---
title: "01 Sort out the results"
author: "Anqi Dai"
date: "2/20/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = F, warning = F)
```

The goal is to do the first two pie charts that illustrate what the AS type distribution is in KD and control respetively, so that the difference of the difference (the significance of the delta PSI) will make more sense.

Keeping the psi that is closer to 0.5 from the 14control and 89control comparison, since the psi will later be used for filtering.

Don't need to filter since we are not caring about the significance level here.

## sorting out whippet results

```{r}
# get the whippet result that has the PSI A and B in all overlapped locus 
# the significance donesn't matter here since we don't care about the truethness of the delta psi

# find the overlap coord between 14 and 89

w14 <- read_tsv('../../03_ASprofile/data/whippetResult/RF14_vs_cntrl_output.diff')%>%
  dplyr::select(Coord, Type, Psi_A, Psi_B) %>%
  filter(!duplicated(Coord))
w89 <- read_tsv('../../03_ASprofile/data/whippetResult/RF89_vs_cntrl_output.diff')%>%
  dplyr::select(Coord, Type_89 = Type, Psi_A_89 = Psi_A) %>%
  filter(!duplicated(Coord))

ol_locus <- intersect(w14$Coord, w89$Coord)

whi <- w14 %>%
  filter(Coord %in% ol_locus) %>%
  inner_join(w89 %>%
              filter(Coord %in% ol_locus) , by = 'Coord')

# make sure they have the same type between the 2
whi %>%
  mutate(same = ifelse(Type == Type_89, T, F)) %>%
  filter(same == F) %>%
  nrow

# filter out the ones that they have different type
whi <- whi %>%
  filter(Type == Type_89)

# now keep the Psi_A that is closer to 0.5
whippet <- whi %>%
  mutate(dis14 = abs(Psi_A - 0.5),
         dis89 = abs(Psi_A_89 - 0.5)) %>%
  mutate(keep_A =  ifelse(dis14 < dis89,Psi_A, Psi_A_89 ))

# the final table for whippet
final_whippet <- whippet %>%
  dplyr::select(locus = Coord, type = Type, keep_A, Psi_B) %>%
  rename(Psi_A = keep_A)
```

## sorting out rmats results

```{r}
# now do the above for the rMATS result

# The JCEC.txt files of the AS types
path <- '../../06_rMATS/data/rMATS_result/RF14_control'
fns <- file.path(path, list.files(path, 'JCEC.txt'))

# A3SS 
A3SS <- read_tsv(fns[1]) %>%
  dplyr::select(chr, longExonStart_0base, shortES, geneSymbol, FDR, IncLevel1, IncLevel2) %>%
  mutate(Significance = ifelse(FDR < 0.05, TRUE, FALSE)) %>%
  rename(start = longExonStart_0base, end = shortES, external_gene_name = geneSymbol, Psi_A = IncLevel1, Psi_B = IncLevel2) %>%
  separate(Psi_A, into = c('Psi_A1','Psi_A2','Psi_A3'), sep = ',', convert = T) %>%
  separate(Psi_B, into = c('Psi_B1','Psi_B2','Psi_B3'), sep = ',', convert = T) %>%
  mutate(Psi_A = mean(c(Psi_A1, Psi_A2, Psi_A3), na.rm = T),
         Psi_B = mean(c(Psi_B1, Psi_B2, Psi_B3), na.rm = T))  %>%
  mutate(start = start + 1) %>%
  mutate(locus = paste(chr, paste(start,end, sep='-'),sep=':')) %>%
  mutate(tool = 'rMATS',
         type = 'AA') %>%
  dplyr::select(external_gene_name, Significance:type)

# A5SS 
A5SS <- read_tsv(fns[2]) %>%
  dplyr::select(chr, shortEE, longExonEnd, geneSymbol, FDR, IncLevel1, IncLevel2) %>%
  mutate(Significance = ifelse(FDR < 0.05, TRUE, FALSE)) %>%
  rename(start = shortEE, end = longExonEnd, external_gene_name = geneSymbol,  Psi_A = IncLevel1, Psi_B = IncLevel2) %>%
  separate(Psi_A, into = c('Psi_A1','Psi_A2','Psi_A3'), sep = ',', convert = T) %>%
  separate(Psi_B, into = c('Psi_B1','Psi_B2','Psi_B3'), sep = ',', convert = T) %>%
  mutate(Psi_A = mean(c(Psi_A1, Psi_A2, Psi_A3), na.rm = T),
         Psi_B = mean(c(Psi_B1, Psi_B2, Psi_B3), na.rm = T))  %>%
  mutate(locus = paste(chr, paste(start,end, sep='-'),sep=':')) %>%
  mutate(tool = 'rMATS',
         type = 'AD') %>%
  dplyr::select(external_gene_name, Significance:type)
  
# RI 
RI <- read_tsv(fns[4]) %>%
  dplyr::select(chr, upstreamEE, downstreamES, geneSymbol, FDR, IncLevel1, IncLevel2) %>%
  mutate(Significance = ifelse(FDR < 0.05, TRUE, FALSE)) %>%
  rename(start = upstreamEE, end = downstreamES, external_gene_name = geneSymbol, Psi_A = IncLevel1, Psi_B = IncLevel2) %>%
  separate(Psi_A, into = c('Psi_A1','Psi_A2','Psi_A3'), sep = ',', convert = T) %>%
  separate(Psi_B, into = c('Psi_B1','Psi_B2','Psi_B3'), sep = ',', convert = T) %>%
  mutate(Psi_A = mean(c(Psi_A1, Psi_A2, Psi_A3), na.rm = T),
         Psi_B = mean(c(Psi_B1, Psi_B2, Psi_B3), na.rm = T))  %>%
  mutate(locus = paste(chr, paste(start,end, sep='-'),sep=':')) %>%
  mutate(tool = 'rMATS',
         type = 'RI') %>%
  dplyr::select(external_gene_name, Significance:type)

# SE
SE <-  read_tsv(fns[5]) %>%
  dplyr::select(chr, exonStart_0base, exonEnd, geneSymbol, FDR, IncLevel1, IncLevel2) %>%
  mutate(Significance = ifelse(FDR < 0.05, TRUE, FALSE)) %>%
  rename(start = exonStart_0base, end = exonEnd, external_gene_name = geneSymbol, Psi_A = IncLevel1, Psi_B = IncLevel2) %>%
  separate(Psi_A, into = c('Psi_A1','Psi_A2','Psi_A3'), sep = ',', convert = T) %>%
  separate(Psi_B, into = c('Psi_B1','Psi_B2','Psi_B3'), sep = ',', convert = T) %>%
  mutate(Psi_A = mean(c(Psi_A1, Psi_A2, Psi_A3), na.rm = T),
         Psi_B = mean(c(Psi_B1, Psi_B2, Psi_B3), na.rm = T))  %>%
  mutate(start = start + 1) %>%
  mutate(locus = paste(chr, paste(start,end, sep='-'),sep=':')) %>%
  mutate(tool = 'rMATS',
         type = 'CE') %>%
  dplyr::select(external_gene_name, Significance:type)

# MXE
MXE <- read_tsv(fns[3]) %>%
  dplyr::select(chr, upstreamES, downstreamEE, geneSymbol, FDR, IncLevel1, IncLevel2) %>%
  mutate(Significance = ifelse(FDR < 0.05, TRUE, FALSE)) %>%
  rename(start = upstreamES, end = downstreamEE, external_gene_name = geneSymbol, Psi_A = IncLevel1, Psi_B = IncLevel2) %>%
  separate(Psi_A, into = c('Psi_A1','Psi_A2','Psi_A3'), sep = ',', convert = T) %>%
  separate(Psi_B, into = c('Psi_B1','Psi_B2','Psi_B3'), sep = ',', convert = T) %>%
  mutate(Psi_A = mean(c(Psi_A1, Psi_A2, Psi_A3), na.rm = T),
         Psi_B = mean(c(Psi_B1, Psi_B2, Psi_B3), na.rm = T))  %>%
  mutate(locus = paste(chr, paste(start,end, sep='-'),sep=':')) %>%
  mutate(tool = 'rMATS',
         type = 'MXE') %>%
  dplyr::select(external_gene_name, Significance:type)
```

```{r}
df14 <- bind_rows(A3SS, A5SS, RI, SE, MXE) %>%
  filter(!duplicated(locus)) %>%
  dplyr::select(locus, type, Psi_A, Psi_B)
```

```{r}
# The JCEC.txt files of the AS types
path <- '../../06_rMATS/data/rMATS_result/RF89_control'
fns <- file.path(path, list.files(path, 'JCEC.txt'))

# A3SS 
A3SS <- read_tsv(fns[1]) %>%
  dplyr::select(chr, longExonStart_0base, shortES, geneSymbol, FDR, IncLevel1, IncLevel2) %>%
  mutate(Significance = ifelse(FDR < 0.05, TRUE, FALSE)) %>%
  rename(start = longExonStart_0base, end = shortES, external_gene_name = geneSymbol, Psi_A = IncLevel1, Psi_B = IncLevel2) %>%
  separate(Psi_A, into = c('Psi_A1','Psi_A2','Psi_A3'), sep = ',', convert = T) %>%
  separate(Psi_B, into = c('Psi_B1','Psi_B2','Psi_B3'), sep = ',', convert = T) %>%
  mutate(Psi_A = mean(c(Psi_A1, Psi_A2, Psi_A3), na.rm = T),
         Psi_B = mean(c(Psi_B1, Psi_B2, Psi_B3), na.rm = T))  %>%
  mutate(start = start + 1) %>%
  mutate(locus = paste(chr, paste(start,end, sep='-'),sep=':')) %>%
  mutate(tool = 'rMATS',
         type = 'AA') %>%
  dplyr::select(external_gene_name, Significance:type)

# A5SS 
A5SS <- read_tsv(fns[2]) %>%
  dplyr::select(chr, shortEE, longExonEnd, geneSymbol, FDR, IncLevel1, IncLevel2) %>%
  mutate(Significance = ifelse(FDR < 0.05, TRUE, FALSE)) %>%
  rename(start = shortEE, end = longExonEnd, external_gene_name = geneSymbol,  Psi_A = IncLevel1, Psi_B = IncLevel2) %>%
  separate(Psi_A, into = c('Psi_A1','Psi_A2','Psi_A3'), sep = ',', convert = T) %>%
  separate(Psi_B, into = c('Psi_B1','Psi_B2','Psi_B3'), sep = ',', convert = T) %>%
  mutate(Psi_A = mean(c(Psi_A1, Psi_A2, Psi_A3), na.rm = T),
         Psi_B = mean(c(Psi_B1, Psi_B2, Psi_B3), na.rm = T))  %>%
  mutate(locus = paste(chr, paste(start,end, sep='-'),sep=':')) %>%
  mutate(tool = 'rMATS',
         type = 'AD') %>%
  dplyr::select(external_gene_name, Significance:type)
  
# RI 
RI <- read_tsv(fns[4]) %>%
  dplyr::select(chr, upstreamEE, downstreamES, geneSymbol, FDR, IncLevel1, IncLevel2) %>%
  mutate(Significance = ifelse(FDR < 0.05, TRUE, FALSE)) %>%
  rename(start = upstreamEE, end = downstreamES, external_gene_name = geneSymbol, Psi_A = IncLevel1, Psi_B = IncLevel2) %>%
  separate(Psi_A, into = c('Psi_A1','Psi_A2','Psi_A3'), sep = ',', convert = T) %>%
  separate(Psi_B, into = c('Psi_B1','Psi_B2','Psi_B3'), sep = ',', convert = T) %>%
  mutate(Psi_A = mean(c(Psi_A1, Psi_A2, Psi_A3), na.rm = T),
         Psi_B = mean(c(Psi_B1, Psi_B2, Psi_B3), na.rm = T))  %>%
  mutate(locus = paste(chr, paste(start,end, sep='-'),sep=':')) %>%
  mutate(tool = 'rMATS',
         type = 'RI') %>%
  dplyr::select(external_gene_name, Significance:type)

# SE
SE <-  read_tsv(fns[5]) %>%
  dplyr::select(chr, exonStart_0base, exonEnd, geneSymbol, FDR, IncLevel1, IncLevel2) %>%
  mutate(Significance = ifelse(FDR < 0.05, TRUE, FALSE)) %>%
  rename(start = exonStart_0base, end = exonEnd, external_gene_name = geneSymbol, Psi_A = IncLevel1, Psi_B = IncLevel2) %>%
  separate(Psi_A, into = c('Psi_A1','Psi_A2','Psi_A3'), sep = ',', convert = T) %>%
  separate(Psi_B, into = c('Psi_B1','Psi_B2','Psi_B3'), sep = ',', convert = T) %>%
  mutate(Psi_A = mean(c(Psi_A1, Psi_A2, Psi_A3), na.rm = T),
         Psi_B = mean(c(Psi_B1, Psi_B2, Psi_B3), na.rm = T))  %>%
  mutate(start = start + 1) %>%
  mutate(locus = paste(chr, paste(start,end, sep='-'),sep=':')) %>%
  mutate(tool = 'rMATS',
         type = 'CE') %>%
  dplyr::select(external_gene_name, Significance:type)

# MXE
MXE <- read_tsv(fns[3]) %>%
  dplyr::select(chr, upstreamES, downstreamEE, geneSymbol, FDR, IncLevel1, IncLevel2) %>%
  mutate(Significance = ifelse(FDR < 0.05, TRUE, FALSE)) %>%
  rename(start = upstreamES, end = downstreamEE, external_gene_name = geneSymbol, Psi_A = IncLevel1, Psi_B = IncLevel2) %>%
  separate(Psi_A, into = c('Psi_A1','Psi_A2','Psi_A3'), sep = ',', convert = T) %>%
  separate(Psi_B, into = c('Psi_B1','Psi_B2','Psi_B3'), sep = ',', convert = T) %>%
  mutate(Psi_A = mean(c(Psi_A1, Psi_A2, Psi_A3), na.rm = T),
         Psi_B = mean(c(Psi_B1, Psi_B2, Psi_B3), na.rm = T))  %>%
  mutate(locus = paste(chr, paste(start,end, sep='-'),sep=':')) %>%
  mutate(tool = 'rMATS',
         type = 'MXE') %>%
  dplyr::select(external_gene_name, Significance:type)
```

```{r}
df89 <- bind_rows(A3SS, A5SS, RI, SE, MXE) %>%
  filter(!duplicated(locus))  %>%
  dplyr::select(locus, type_89 = type, Psi_A_89 = Psi_A, Psi_B_89 = Psi_B)

rmats <- df14 %>%
  inner_join(df89, by  = 'locus') %>%
  filter(type == type_89) %>%
  mutate(dis14_A = abs(Psi_A - 0.5),
         dis89_A = abs(Psi_A_89 - 0.5),
         dis14_B = abs(Psi_B - 0.5),
         dis89_B = abs(Psi_B_89 - 0.5)) %>%
  mutate(keep_A =  ifelse(dis14_A < dis89_A,Psi_A, Psi_A_89 ),
         keep_B =  ifelse(dis14_B < dis89_B,Psi_B, Psi_B_89 )) %>%
  dplyr::select(locus, type, keep_A, keep_B) %>%
  rename(Psi_A = keep_A, Psi_B = keep_B)

```

## deduplicate between rmats and whippet

```{r}
# the duplicated locus
dup_locus <- intersect(rmats$locus, final_whippet$locus)

#  join the unduplicated part in the two
undup <- bind_rows(
  rmats %>%
    filter(! locus %in% dup_locus),
  final_whippet %>%
    filter(! locus %in% dup_locus)
)

# keep the closer to 0.5 psi in the duplicated locus
rmats_dup <- rmats %>%
  filter(locus %in% dup_locus)

whippet_dup <- final_whippet %>%
  filter(locus %in% dup_locus) %>%
  rename(type_w = type, Psi_A_w = Psi_A,  Psi_B_w = Psi_B)

dups <- rmats_dup %>%
  inner_join(whippet_dup, by = 'locus') %>%
  filter(type == type_w) %>%
  mutate(disW_A = abs(Psi_A_w - 0.5),
         disR_A = abs(Psi_A - 0.5),
         disW_B = abs(Psi_B_w - 0.5),
         disR_B = abs(Psi_B - 0.5)) %>%
  mutate(keep_A =  ifelse(disW_A < disR_A,Psi_A_w, Psi_A ),
         keep_B =  ifelse(disW_B < disR_B,Psi_B_w, Psi_B )) %>%
  dplyr::select(locus, type, keep_A, keep_B) %>%
  rename(Psi_A = keep_A, Psi_B = keep_B)

WR_combined <- bind_rows(dups, undup)   
```

## sorting out the IRfinder results

```{r}
IR14 <- read_tsv('../../05_IRfinder/data/rawIRfinderResult/RF14_vs_ctrl.txt', comment = '#')  %>%
  rename_all( funs(stringr::str_replace_all(., '-', '_'))) %>%
  mutate(type = 'RI')  %>%
  dplyr::select(Chr, Start, End, type, A_IRratio, B_IRratio) %>%
  mutate(Start = Start + 1) %>%
  mutate(locus = paste(Chr, paste(Start, End, sep = '-'), sep = ':')) %>%
  dplyr::select(type:locus) %>%
  filter(!duplicated(locus)) %>%
  rename(Psi_A = A_IRratio, Psi_B = B_IRratio)
 

IR89 <- read_tsv('../../05_IRfinder/data/rawIRfinderResult/RF89_vs_ctrl.txt', comment = '#')  %>%
  rename_all( funs(stringr::str_replace_all(., '-', '_'))) %>%
  mutate(type = 'RI')  %>%
  dplyr::select(Chr, Start, End, type, A_IRratio, B_IRratio)  %>%
  mutate(Start = Start + 1) %>%
  mutate(locus = paste(Chr, paste(Start, End, sep = '-'), sep = ':')) %>%
  filter(!duplicated(locus)) %>%
  dplyr::select(locus, type_89 = type, Psi_A_89 = A_IRratio, Psi_B_89 = B_IRratio)
  
```

```{r}
# join them together
IRfinder <- IR14 %>%
  inner_join(IR89, by  = 'locus') %>%
  mutate(dis14_A = abs(Psi_A - 0.5),
         dis89_A = abs(Psi_A_89 - 0.5)) %>%
  mutate(keep_A =  ifelse(dis14_A < dis89_A,Psi_A, Psi_A_89 )) %>%
  dplyr::select(locus, type, keep_A, Psi_B) %>%
  rename(Psi_A = keep_A)

```

## deduplicate between IRfinder and the other 2

```{r}
# the duplicated locus
dup_locus <- intersect(IRfinder$locus, WR_combined$locus)

#  join the unduplicated part in the two
undup <- bind_rows(
  IRfinder %>%
    filter(! locus %in% dup_locus),
  WR_combined %>%
    filter(! locus %in% dup_locus)
)

# keep the closer to 0.5 psi in the duplicated locus
IRfinder_dup <- IRfinder %>%
  filter(locus %in% dup_locus)

two_dup <- WR_combined %>%
  filter(locus %in% dup_locus) %>%
  rename(type_2 = type, Psi_A_2 = Psi_A,  Psi_B_2 = Psi_B)

# merge the dups together
dups <- IRfinder_dup %>%
  inner_join(two_dup, by = 'locus') %>%
  filter(type == type_2) %>%
  mutate(dis2_A = abs(Psi_A_2 - 0.5),
         disIR_A = abs(Psi_A - 0.5),
         dis2_B = abs(Psi_B_2 - 0.5),
         disIR_B = abs(Psi_B - 0.5)) %>%
  mutate(keep_A =  ifelse(dis2_A < disIR_A,Psi_A_2, Psi_A ),
         keep_B =  ifelse(dis2_B < disIR_B,Psi_B_2, Psi_B )) %>%
  dplyr::select(locus, type, keep_A, keep_B) %>%
  rename(Psi_A = keep_A, Psi_B = keep_B)

# combined to make the final one! Thank god!!!

final <- bind_rows(dups, undup)
```

```{r}
# write out the final data
write_csv(final, '../output/all_three_PSI_A_B_combined_not_filtered.csv')
```

