---
title: "[DAY 3] IRfinder result analysis: 14 VS control and 89 VS control"
author: "Anqi Dai"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 3
    code_folding: hide
---

```{r setup, include=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE, message=FALSE, warning = FALSE)
```


```{r}
library(tidyverse)
library(ggpubr)
```


## Filtering to get the significant results:

* Remove records marked with "known-exon". "known-exon" indicates that there are known transcripts both including and splicing out this region. IRFinder quantifies the relative abundance of these occurrences.
* Remove records marked with "MinorIsoform" either in experiment or control. "MinorIsoform" indicates IRFinder has detected there is additional alternate-splicing occurring, as such the detection of intron-retention may be confounded by other changes in alternate splicing.
* Adjust the P-value with BH method (The Audic and Claverie test is used for low replicates situations (<= 3 replicates). The assumption is that the IR ratio (log transformed) follows a Poisson distribution)
* Select the records with padj < 0.05


```{r}
# load the result table
RES <- read_csv('../data/IRfinder.res.filtered.14n89.analysis2.csv')
```


## Overview of the overlap.


## The analysis result

### Length of the RI

```{r barplot.IRlen}
# the summarised plot of the mean RI_len between the sig and non-sig
# boxplot 
RES %>%
  dplyr::select(Locus, Significance, RI_len, Group) %>%
  spread(key = Group, value = RI_len) %>%
  ggboxplot(x = "Significance", y = c('RF14', 'RF89'), combine = T,
            #add = "mean_se", 
            #fill = 'gray',
            #error.plot = "upper_errorbar",
            ylab = 'Length of retained intron',
            #ylim = c(0, 3500),
            title = 'Len of RI boxplot') + 
  #scale_y_continuous(expand = c(0, 0))  +
  ggsave('../figs/Len of RI in both between sig and non sig_boxplot.jpg', width = 8, height = 8, dpi = 300)

```


### Compare the TPM

#### Histogram

```{r}
# make a faceted histogram of the tpm counts between the sig and non sig
# histogram
RES %>%
  dplyr::select(Locus, Significance, Group, RFExper, RFControl) %>%
  spread(key = Group, value = RFExper) %>%
  gghistogram(x = c('RF14','RF89', 'RFControl'), combine = T,
              palette = 'lancet', 
              color = 'white', 
              fill = 'Significance', 
              bins = 50,
              title = 'Histogram of log10(mean TPM)',
              xlab = 'log10(mean TPM)')  +
  scale_x_log10() +
  ggsave('../figs/Histogram of log10(mean TPM).jpg', width = 12, height = 8, dpi = 300)
```

#### boxplot with statistic test comparing the mean (t test)

```{r}
# boxplot
RES %>%
  dplyr::select(Locus, Significance, Group, RFExper, RFControl) %>%
  spread(key = Group, value = RFExper) %>%
  ggboxplot(x = "Significance", y = c("RF14", 'RF89','RFControl'), combine = T,
            ylim = c(0.5, 10000),
            color = "Significance", palette = 'lancet',
            alpha = 0.3,
            ylab = 'log10(mean TPM)',
            xlab = 'Significance level'
             ) + 
  scale_y_log10() +
  stat_compare_means(comparisons = list(c('TRUE', 'FALSE')), method = 't.test')




```

```{r} 
# it would be interesting to see the TRUE in the 14, 89 and control how would they compare
RES %>%
  dplyr::select(Locus, Significance, Group, RFExper, RFControl) %>%
  spread(key = Group, value = RFExper) %>%
  filter(Significance) %>%
  gather(key = Group, value = MeanTPM, RFControl:RF89) %>%
  mutate(Group = factor(Group, levels = c('RFControl','RF14','RF89'))) %>%
  ggboxplot(x = "Group", y = 'MeanTPM',
            #ylim = c(0.5, 10000),
            #add = "jitter",
            color = "Group", palette = 'lancet',
            title = 'mean TPM comparison in significant RI genes',
            ylab = 'log10(mean TPM)'
            #xlab = 'Significance level'
             ) + 
  scale_y_log10() +
  stat_compare_means(comparisons = list(c('RF14', 'RFControl'), 
                                        c('RF89', 'RFControl'),
                                        c('RF14', 'RF89')), method = "t.test") +
  ggsave('../figs/mean TPM comparison in significant RI genes.png', width = 6, height = 5, dpi = 300)

# a histogram to see the distribution (which is actually higher or lower)

p <- RES %>%
  dplyr::select(Locus, Significance, Group, RFExper, RFControl) %>%
  spread(key = Group, value = RFExper) %>%
  filter(Significance) %>%
  gather(key = Group, value = MeanTPM, RFControl:RF89) %>%
  mutate(Group = factor(Group, levels = c('RFControl','RF14','RF89'))) %>%
  gghistogram(x = 'MeanTPM', 
              palette = 'lancet',
              color = 'white', fill = 'Group', 
              bins = 80,
              add = "median",
              title = 'Histogram of log10(mean TPM)(median shown)',
              xlab = 'log10(mean TPM)')  +
  scale_x_log10() 
  

facet(p, facet.by = "Group", nrow = 3)
```

### Compare the GC content of the RI

```{r}
# histogram

RES %>%
  dplyr::select(Locus, Significance, GC_per, Group) %>%
  spread(key = Group, value = GC_per) %>%
  gghistogram(x = c('RF14','RF89'),
              combine = T,
              palette = 'lancet',
              color = 'white', fill = 'Significance', 
              bins = 40,
              add = "median",
              title = 'Histogram of GC% (median shown)',
              xlab = 'GC% ') 
```

```{r}
# boxplot

RES %>%
  dplyr::select(Locus, Significance, GC_per, Group) %>%
  spread(key = Group, value = GC_per) %>%
  ggboxplot(x = "Significance", y =  c('RF14','RF89'), combine = T,
            #ylim = c(0.5, 10000),
            #add = "jitter",
            color = "Significance", palette = 'lancet',
            xlab = 'Significance',
            title = 'GC% among sig or not sig locus',
            ylab = 'GC%'
             ) + 
  stat_compare_means(comparisons = list(c('TRUE', 'FALSE')), method = 't.test')
```

### The locus of the RI 

Compare the proportion of the middle point of RI at the whole gene length

```{r}
# plot the -log10(Padj) VS RI_prop (dot plot)
# start from the ggplot
Col <- c('#00468B', '#EC0000')

RES %>%
  dplyr::select(Locus, Significance, padj, RI_prop, Group ) %>%
  mutate(log10P = -log10(padj))  %>%
  filter(!is.na(RI_prop) & !is.infinite(log10P)) %>%
  mutate(RI_prop = round(RI_prop, 2)) %>%
  ggplot(aes(x =RI_prop, y = log10P, color = Significance, shape = Significance ))+
  geom_point(alpha = 0.7, size = 2) + 
  labs(x = 'Proportional locus of the RI',
       y = '-log10(padj)' ,
       title = 'log10(padj) VS Proportional locus of the RI') +
  scale_fill_manual(values = Col) +
  scale_color_manual(values = Col)  +
  facet_wrap( ~ Group) + 
  ggsave('../figs/log10(padj) VS Proportional locus of the RI(dot)_both.jpg', width = 10, height = 7, dpi = 300)

```

```{r}
# color the histogram with the factor of the log10P
locus_sig_level <- RES %>%
  dplyr::select(Locus, Significance, padj, RI_prop, Group ) %>%
  mutate(log10P = -log10(padj)) %>%
  filter(!is.na(RI_prop) & !is.infinite(log10P)) %>%
  mutate(neg_log10P = ifelse(is.infinite(log10P), 'Inf',
                            ifelse(log10P > 2^6, '>64',
                                   ifelse(log10P > 2^4, '>16',                                                                         ifelse(log10P > 2^2, '>4',                                                                           ifelse(log10P > 2^0, '>1', '<1')))))) %>%
  mutate(neg_log10P = factor(neg_log10P, levels = c('<1', '>1', '>4', '>16', '>64')))


# histogram
library(RColorBrewer)
col_palette <- brewer.pal(n = 5, name = "Dark2") 

locus_sig_level %>%
  mutate(RI_prop = round(RI_prop, 2)) %>%
  ggplot(aes(x = RI_prop, fill = neg_log10P))+
  geom_histogram( ) + 
  labs(x = 'Proportional locus of the RI',
       y = '-log10(padj)' ,
       title = 'log10(padj) VS Proportional locus of the RI') +
  scale_fill_manual(values = col_palette) +
  scale_color_manual(values = col_palette)  +
  facet_wrap(Group ~ Significance) +
  ggsave('../figs/log10(padj) VS Proportional locus of the RI(histogram_color)_both.jpg', width = 10, height = 10, dpi = 300)

```

```{r}
# boxplot

locus_sig_level %>%
  dplyr::select(Locus, Significance, RI_prop, Group) %>%
    filter(!is.na(RI_prop)) %>%
  spread(key = Group, value = RI_prop) %>%
  ggboxplot(x = 'Significance', y = c('RF14','RF89'),combine = T,
            color = "Significance", palette = 'lancet',
            #add = 'jitter',
            title ='Locus of the RI in proportion to the gene length') +
  stat_compare_means(comparisons = list(c('FALSE', 'TRUE'))) 



```


### Strength of the splice sites

```{r}
RES %>%
  dplyr::select(Locus, Significance, Splicesite_pair, Group ) %>%
  ggplot(aes(Splicesite_pair)) +
  geom_bar() +
  geom_text(aes(label=..count..), stat='count',  vjust=-1) +
  ylim(0,1700) +
  facet_wrap(Group ~ Significance) +
  labs(title = 'Barplot of the splice site type among sig and non-sig genes')
  
```

* Score 5' splice site: Each sequence must be 9 bases long. 3 bases in exon and 6 bases in intron
* Score 3' splice site: Each sequence must be 23 bases long. 20 bases in the intron and 3 base in the exon


```{r score.plot}
# boxplot

# ss5

RES %>%
  dplyr::select(Locus, Significance, ss5_score , Group ) %>%
  spread(key = Group, value = ss5_score) %>%
  ggboxplot(x = 'Significance', y = c('RF14','RF89'),combine = T,
            add = 'jitter', add.params = list(alpha = 0.15, fill = 'Significance'),
            title = 'boxplot of the distribution of the ss5 MaxEnt score',
            palette = 'lancet', color = 'Significance')   +
  stat_compare_means(comparisons = list(c('TRUE', 'FALSE')))


# ss3

RES %>%
  dplyr::select(Locus, Significance, ss3_score , Group ) %>%
  spread(key = Group, value = ss3_score) %>%
  ggboxplot(x = 'Significance', y = c('RF14','RF89'),combine = T,
            add = 'jitter', add.params = list(alpha = 0.15, fill = 'Significance'),
            title = 'boxplot of the distribution of the ss3 MaxEnt score',
            palette = 'lancet', color = 'Significance')   +
  stat_compare_means(comparisons = list(c('TRUE', 'FALSE')))


```




